Topological Equivariant Artist Model

### 2021-03-26 slides/main.tex
lets not delete my appendix

\usepackage{appendixnumberbeamer} 
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt,
                   autogobble, 
                   frame=lines,
                   bgcolor=bg,
                   framesep=3mm} 
\usepackage{notation} % move this later

@@

\begin{frame}{Visualizations are structure preserving maps}
    \begin{figure}
        \includegraphics[width=\textwidth]{figures/intro/viz_same.png}
    \end{figure}
    \begin{description}
        \item[equivariance] properties of data and visual encoding match
        \item[continuity] connectivity of data and visual encoding match
        \item[composibility] structure preserved by individual components is preserved in combined components
    \end{description}
\end{frame}

\begin{frame}{Tools are tuned to continuity \cite{HeerSoftware2006}}
\begin{table}
    %\renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|L|l|l|}\toprule
        \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
        ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
        Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
        Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
         Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} & &\\
         \bottomrule
    \end{tabulary}
\end{table}
\end{frame}

\begin{frame}{Visualizations are tuned to continuinity\cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.5\textheight]{figures/intro/dataset_diagram.png}
        \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
    \end{figure}
    \begin{enumerate}
        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$Seaborn\cite{waskom2020seaborn}, xarray \cite{hoyer2017xarray} 
        \item D3 \cite{bostockDataDrivenDocuments2011}
        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveci2012vtk}(MayaVi\cite{RamachandranMayaVI2011})$\rightarrow$ Titan\cite{brianwylieUnifiedToolkitInformation2009}, ParaView\cite{ahrens2005paraview}
    \end{enumerate}
\end{frame}

\begin{frame}{Tools are evaluated on the structure they preserve}
    \begin{description}
        \item[Naturalness] easier to understand when properties match \cite{norman_things_smart}
        \item[Graphical Integrity] graphs show \textbf{only} the data \cite{tufteVisualDisplayQuantitative2001}
        \item[Expressiveness] structure preserving mappings from data to graphic \cite{mackinlayAutomatingDesignGraphical1986}
        \item[Effectiveness] design choices made in deference to perceptual saliency \cite{clevelandResearchStatisticalGraphics1987,clevelandGraphicalPerceptionTheory1984,chambersGraphicalMethodsData1983a, munznerVisualizationAnalysisDesign2014}
    \end{description}
\end{frame}

\subsection{Contributions}
\begin{frame}{Contributions}
    \begin{description}
        \item[Topological] topology preserving relationship between data and graphic via continuous maps
        \item[Equivariant] property preservation from data component to visual representation as equivariant maps that carry a homomorphism of monoid actions
        \item[Artist] functional oriented visualization tool architecture built on the mathematical model to demonstrate the utility of the model
        \item [Model] prototype of the architecture built on Matplotlib's infrastructure to demonstrate the feasibility of the model
    \end{description}
\end{frame}

@@

\begin{frame}{Topological Equivariant Artist Model}
    \centering
    An artist $\mathscr{\vartist}$ is an equivariant map 
    \begin{equation}\Huge
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation}
    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space.

\end{frame}

\subsection{Data Model}

\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
    \begin{equation}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation}
    \begin{description}
        \item[base space \dbase] continuity
        \item[fiber space \dfiber] variable characteristics
        \item[total space \dtotal] dataset topology
    \end{description}
\end{frame}

\begin{frame}{Encode variable types in a schema like fiber \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
    \begin{figure}[H]
        \centering
        \includegraphics[height=.3\textwidth]{figures/math/fiber.png}
        \label{fig:data_fiber_example}
    \end{figure}
    \begin{block}{Monoid actions define the structure on $\dfiber_i$}
        \begin{equation}
            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
        \end{equation}

        $\monoid_{i}$ actions (e.g. rotation, partial ordering) 
        \begin{equation}
            \bullet: \monoid_i\times \dfiber_i \rightarrow \dfiber_i
        \end{equation}
        must be associative and have an identity element.

    \end{block}
\end{frame}

\begin{frame}{Structural \textit{keys} with associated \textit{values} \cite{munznerVisualizationAnalysisDesign2014}}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/base.png}
        \label{fig:base_example}
    \end{figure}

    \begin{block}{\dbase\ is an indexing space independent of component semantics}
        \begin{equation}
            \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
        \end{equation}
    \end{block}
\end{frame}

\begin{frame}{Data are sections \dsection\ on \dtotal}
    For any fiber bundle, there exists a map
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[d, "\pi"'] \\
                              \& \dbase \arrow[u, "\dsection"', bend right]
        \end{tikzcd}
    \end{equation}
     s.t. $\pi(\dsection(\dbasepoint)) = \dbasepoint$. $\Gamma(\dtotal)$ is the set of all global sections.
     \begin{figure}[H]
        \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
        \label{fig:data_sections}
    \end{figure}
\end{frame}

\subsection{Graphics}
\begin{frame}{Graphic Bundle $(\gtotal,\,\gbase,\,\pi ,\,\gfiber)$}
    Continuity is preserved via the deformation retraction $\vindex: \gbase\rightarrow\dbase$
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']                                 \\
                              \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[ll, "\xi"'] \arrow[u, "\rho"', bend right]
            \end{tikzcd}
        \end{equation}
    \gfiber\ is a proxy for the target display, for example $(x,\, y,\, r,\, g,\, b) \in \gfiber$
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/retraction_maps.png}
        \label{fig:graphic_retraction_map}
    \end{figure}
\end{frame}

\subsection{Artist}
\begin{frame}{Topological Equivariant Artist Model}
The topological artist is a sheaf map
\begin{equation}
\vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation}
that carries homomorphism of monoid actions $\varphi: \monoid \rightarrow \monoid^{\prime}$ \cite{cegarraCohomologyMonoidsOperators2019} 
\begin{equation}
\vartist(m\cdot \delement) = \varphi(m)\cdot \vartist(\delement) 
\end{equation} 
and is implemented via an intermediate visual bundle $(\vtotal,\,\dbase,\,\pi ,\,\vfiber)$
\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}[ampersand replacement=\&]
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] \& \vtotal \arrow[d, "\pi"] \& \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] \& \gtotal \arrow[ld, "\pi"] \\
                                              \& \dbase                  \& \gbase \arrow[l, "\vindex"']                                              \&                    
        \end{tikzcd}
\end{equation}
\end{frame}

\begin{frame}{Existing models don't have continuity, many evaluative}
        \begin{description}
            \item[algebraic process] data and viz transforms are symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
            \begin{equation*}
                \begin{tikzcd}[ampersand replacement=\&]
                    \dtotal \arrow[d, "m"'] \arrow[r, "r_1"] \& \vtotal \arrow[r, "\nu"]  \& \gtotal \arrow[d, "\varphi(m)"] \\
                    \dtotal \arrow[r, "r_2"']                     \& \vtotal \arrow[r, "\nu"'] \& \gtotal                  
                \end{tikzcd}
            \end{equation*}

            \item[language] APT and GoG: syntax, semantics, and grammar\\ 
            \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}
            \item[functional dependencies] relationship between components \cite{sugibuchiFramwork2009}
            \item[category theory]  \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}

        \end{description}
\end{frame}

\begin{frame}{Visual Channel Encoders}
We define the visual transformers \vchannel\ on components of the data bundle $\dsection_{i}$
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps with the constraint 
\begin{equation}
    \vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\end{equation} 
where $\varphi:\monoid\rightarrow \monoid^{\prime}$ carries a homomorphism of monoid actions. 
\end{frame}

\begin{frame}{Equivariance: Partial Order \& and Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}

\begin{table}[H]
    \begin{tabularx}{\textwidth}{|l|l|X|}\toprule
        \textbf{scale} & \textbf{group} & \textbf{constraint} \\\midrule
        nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
        ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
        interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
        ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\ \bottomrule
    \end{tabularx}
\end{table}
\end{frame}

\begin{frame}{Visualization Assembly Function}
    \begin{figure}
        \includegraphics[width=\textwidth]{figures/math/path_of_q.png}
    \end{figure}
\end{frame}

\begin{frame}{Visualization Equivariance}
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \end{figure}
    \begin{prop}
        If for elements of the monoid $m \in \monoid$ and for all $\vsection, \vsection^{\prime} \in X$, we define the monoid action on $X$ so that it is by definition equivariant
        \begin{equation}
        \vmark(\vsection) = \vmark(\vsection^{\prime})\implies \vmark(m\circ\vsection) = \vmark(m\circ\vsection^{\prime})
        \label{eq:math:artist:q}
        \end{equation}
        then a monoid action on $Y$ can be defined as $m\circ \gsection = \gsection^{\prime}$. If and only if \vmark\ satistfies \autoref{eq:math:artist:q}, we can state that the transformed graphic $\gsection^{\prime}=\vmark(m\circ \vsection)$ is equivariant to a monoid action applied on \vmark\ with input $\vsection \in \vmark^{-1}(\gsection)$ that must generate valid \gsection. 
    \end{prop}
\end{frame}

@@

\begin{frame}{Line: $\vmark(xpos, \hat{n_{1}}, ypos, \hat{n_{2}})(\alpha, \beta)$ }
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/line.png}
    \end{figure}
        \begin{equation*}
            \lvert n \rvert = \sqrt{{n_{1}}^2 + {n_{2}}^2},\; 
            \hat{n_{1}} = \frac{n_1}{\lvert n \rvert}, \; \hat{n_{2}} = \frac{n_2}{\lvert n \rvert}
        \end{equation*}
    \begin{align*}
     x = xpos(\vindex(\alpha)) &+ width*\beta\hat{n_1}(\vindex(\alpha)) \\
     y = ypos(\vindex(\alpha)) &+ width*\beta\hat{n_2}(\vindex(\alpha)) 
    \end{align*}
\end{frame}

\begin{frame}{Image $\vmark(xpos, ypos, color)$}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/image.png}
    \end{figure}
    \begin{align*}
        R &= R(\vindex(\alpha, \beta))\\
        G &= G(\vindex(\alpha, \beta))\\
        B &= B(\vindex(\alpha, \beta))
    \end{align*}
\end{frame}

\begin{frame}{Assembly Function Factory}
\begin{equation}\label{eq:qhat_q_s}
    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
\end{equation} 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
\end{figure}
such \gbasepoint\ can be factored out when $\vpreimg = \gbasepoint$
\end{frame}

\begin{frame}{Composition of artists}
    Given the family of artists $(\dtotal_i: i\in I)$ on the same image
    \begin{equation}
        + \coloneqq \underset{i\in I}{\sqcup} \dtotal_{i}
    \end{equation}
    the + operator defines a simple composition of artists.

    When artists share a base space 
    \begin{equation}
        \dbase_2 \hookrightarrow \dbase_1
    \end{equation}
    a composition operator can be defined such that the the artists can be considered to be acting on different components of the same section. 
\end{frame}

\begin{frame}{Equivalance class of artists}
    An approximation of the equivalence class of artists $\vartist^{\prime}$
    \begin{equation}
    \vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}
    \end{equation}
    roughly treats two artists as equivalent if they 
    \begin{itemize}
        \item act on the same visual bundle \vtotal
        \item have the same assembly function \vmark 
        \item have the same continuity map \vindex 
    \end{itemize}
\end{frame}
\section{Prototype}

\begin{frame}{TEAM driven rearchitecture of Matplotlib}
    \begin{itemize}
        \item fiber bundle with schema is a structure rich abstraction
        \begin{itemize}
            \item topologically complex heterogenous data 
            \item target display spaces
        \end{itemize}
        \item structure preserving maps from data to visual:
            \begin{itemize}
                \item data and graphics have equivalent continuity
                \item properties are equivariant under monoid actions
            \end{itemize}  
        \item composable into complex visualizations
    \end{itemize}
\end{frame}

\begin{frame}[fragile]{Artists: Scatter \& Line}
    \begin{figure}[H]
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/line_1.png}
        \end{subfigure}
    \end{figure}
    \begin{columns}
    \column{0.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Point(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \column{.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Line(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \end{columns}
\end{frame}

\subsection{Visual Transforms}
\begin{frame}[fragile]{Visual Transformations}
    \begin{minted}{python}
    cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
    transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
                    'y': {'name': 'v2', 'encoder': lambda x: x},
                    'facecolors': {'name':'v3', 'encoder': cmap}, 
                    's':{'name': None , 
                        'encoder': lambda _: itertools.repeat(.02)}}
    \end{minted}
        \begin{itemize}
            \item \mintinline{python}{lambda x: x} is identity \vchannel\
            \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
            \item \mintinline{python}{color.Categorical} is custom \vchannel
        \end{itemize}
    \end{frame}

    \begin{frame}[fragile]{Complex \vchannel}
    \begin{minted}{python}
    class Categorical:
        def __init__(self, mapping):
            # check that the conversion is to valid colors
            assert(mcolors.is_color_like(color) for color in mapping.values())
            self._mapping = mapping

>>>>

\usepackage{appendixnumberbeamer} 
\usepackage{comment}
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt,
                   autogobble, 
                   frame=lines,
                   bgcolor=bg,
                   framesep=3mm} 
\usepackage{notation} % move this later

@@

\begin{frame}{Visualizations are structure preserving maps}
    %add arrow between data and vizualizations
    % we're focused on building good arrows 
    \begin{figure}
        \includegraphics[width=\textwidth]{figures/intro/viz_same.png}
    \end{figure}
    \pause 
    \begin{description}
        \item[equivariance] properties of data and visual encoding match
    \pause 
        \item[continuity] connectivity of data and visual encoding match
    \end{description}
\end{frame}

\begin{frame}{Domain specific libraries assume data structure\cite{HeerSoftware2006}}
    \begin{table}
        %\renewcommand{\arraystretch}{2}
        \begin{tabulary}{\textwidth}{|L|l|l|}\toprule
            \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
            Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
            Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
             Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} & &\\
             \bottomrule
        \end{tabulary}
    \end{table}
\end{frame}

\begin{frame}{General purpose libraries can't\cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.5\textheight]{figures/intro/dataset_diagram.png}
        \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
    \end{figure}
    \begin{enumerate}
        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$Seaborn\cite{waskom2020seaborn}, xarray \cite{hoyer2017xarray} 
        \item D3 \cite{bostockDataDrivenDocuments2011}
        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveci2012vtk}(MayaVi\cite{RamachandranMayaVI2011})$\rightarrow$ Titan\cite{brianwylieUnifiedToolkitInformation2009}, ParaView\cite{ahrens2005paraview}
    \end{enumerate}
\end{frame}

\begin{frame}{Best practices in visualization design}
    \begin{description}
        \item[Expressiveness] structure preserving mappings\cite{mackinlayAutomatingDesignGraphical1986}
        \pause
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
        \pause
       \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
    \end{description}
\end{frame}

\subsection{Contributions}
\begin{frame}{Contributions}
    \begin{description}
        \item[Topological] continuity  
        \item[Equivariant] monoid action 
        \item[Artist] Matplotlib $\textit{artist}: \textit{data} \rightarrow \textit{graphic}$ 
        \item [Model] 
    \end{description}
\end{frame}

@@

\begin{frame}{Topological Equivariant Artist Model}
    \centering
    An artist $\mathscr{\vartist}$ is an equivariant map 
    \begin{equation*}\Huge
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation*}
    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space. 
\end{frame}

\subsection{Data Model}
\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
    \begin{equation*}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation*}
    \begin{description}
        \pause
        \item[total space \dtotal] dataset topology
        \pause
        \item[fiber space \dfiber] variable schema
        \pause
        \item[base space \dbase] continuity
    \end{description}
\end{frame}

\begin{frame}{Encode variable types in a schema like fiber \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
    \begin{figure}[H]
        \centering
        \includegraphics[width=\textwidth]{figures/math/fiber.png}
    \label{fig:data_fiber_example}
    \end{figure}
\end{frame}

\begin{frame}{Structure of components of \dfiber}
    \begin{figure}[H]
        \centering
        \includegraphics[height=.3\textheight]{figures/math/fiber.png}
    \label{fig:data_fiber_example}
    \end{figure}
        \begin{equation*}
            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
        \end{equation*}
        \pause
        Monoid actions $\monoid_{i}$ (e.g. rotation, partial ordering) define the structure on $\dfiber_i$
        \begin{equation*}
            \bullet: \monoid_i\times \dfiber_i \rightarrow \dfiber_i
        \end{equation*}
        They must be associative and have an identity element. 
\end{frame}

\begin{frame}{\dbase\ is an indexing space}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/base.png}
    \label{fig:base_example}
    \end{figure}
\end{frame}

\begin{frame}{Structural \textit{keys} with associated \textit{values} \cite{munznerVisualizationAnalysisDesign2014}}
\begin{figure}
    \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
    \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
\end{figure}
\pause
\begin{figure}[H]
    \includegraphics[width=.5\textheight]{figures/math/base.png}
    \label{fig:base_example}
\end{figure}
\end{frame}

\begin{frame}{Data are sections \dsection\ on \dtotal}
    For any fiber bundle, there exists a map
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[d, "\pi"'] \\
                              \& \dbase \arrow[u, "\dsection"', bend right]
        \end{tikzcd}
    \end{equation*}
     s.t. $\pi(\dsection(\dbasepoint)) = \dbasepoint$. $\Gamma(\dtotal)$ is the set of all global sections.
     \begin{figure}[H]
        \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
        \label{fig:data_sections}
    \end{figure}
\end{frame}

\subsection{Graphics}
\begin{frame}{Graphic Bundle $(\gtotal,\,\gbase,\,\pi ,\,\gfiber)$} % watch dr explaination
    Continuity is preserved via the many \gbasepoint\ to one \dbasepoint\ map $\vindex: \gbase\rightarrow\dbase$
    \begin{equation*} 
        \begin{tikzcd}[ampersand replacement=\&]
            F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']                                 \\
                              \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[ll, "\xi"'] \arrow[u, "\rho"', bend right]
            \end{tikzcd}
        \end{equation*}
    \pause
    \gfiber\ is a proxy for the target display, for example $(x,\, y,\, r,\, g,\, b) \in \gfiber$
    \pause
    \begin{figure}[H] %%%bring up one at a time
        \includegraphics[width=1\textwidth]{figures/math/retraction_maps.png}
        \label{fig:graphic_retraction_map}
    \end{figure}
\end{frame}

\subsection{Artist}%formalize the structure
\begin{frame}{Visual bundle $(\vtotal,\,\dbase,\,\pi ,\,\vfiber)$}
\begin{equation*}
    \label{eq:artist_diagram}
    \begin{tikzcd}[ampersand replacement=\&]
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] \& \vtotal \arrow[d, "\pi"] \& \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] \& \gtotal \arrow[ld, "\pi"] \\
                                              \& \dbase                  \& \gbase \arrow[l, "\vindex"']                                              \&                    
        \end{tikzcd}
\end{equation*}
\pause
\begin{equation*}
    \mathcal{A}:\mathcal{E}\rightarrow\mathcal{E}
\end{equation*}
\pause
The topological artist is a sheaf map
\begin{equation*}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation*}
\end{frame}

\begin{frame}{Visualization Assembly Function}
    \begin{figure}
        \includegraphics[width=\textwidth]{figures/math/path_of_q.png}
    \end{figure}
    \pause
    \begin{equation*}
        \label{eq:nu_expanded}
        \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
    \end{equation*}
    \pause
    \begin{equation*}
        Q = \nu \circ \tau
    \end{equation*}
\end{frame}

\begin{frame}{Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
\begin{table}[H]
    \begin{tabularx}{\textwidth}{|l|l|X|}\toprule
        \textbf{scale} & \textbf{group} & \textbf{constraint} \\\midrule
        nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
        ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
        interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
        ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\ \bottomrule
    \end{tabularx}
\end{table}
\end{frame}

\begin{frame}{Equivariance: Partial Order}
    \begin{figure} %break out steps/reveal
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}
\end{frame}

\begin{frame}{Visualization Equivariance}
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \end{figure}

    %put equation w/ rho, label rho, rho'
\end{frame}

@@

\begin{frame}{Line: $\vmark(xpos, \hat{n_{1}}, ypos, \hat{n_{2}})(\alpha, \beta)$ }
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/line.png}
    \end{figure}
\end{frame}

\begin{frame}{Image $\vmark(xpos, ypos, color)$}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/image.png}
    \end{figure}
\end{frame}

\begin{frame}{Build \vmark over \dbase:\vmarkd}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
\end{figure}
\begin{equation*}\label{eq:qhat_q_s}
    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
\end{equation*} 
such \gbasepoint\ can be factored out when $\vpreimg = \gbasepoint$
\end{frame}

\begin{frame}{Composition of artists} 
    % make disjointed figure or scatter + line
    Given the family of artists $(\dtotal_i: i\in I)$ on the same image
    \begin{equation*}
        + \coloneqq \underset{i\in I}{\sqcup} \dtotal_{i}
    \end{equation*}
    the + operator defines a simple composition of artists. 
\end{frame}

\section{Prototype}
\begin{frame}{TEAM driven rearchitecture of Matplotlib}
    \begin{itemize}
        \begin{itemize}
            \item topologically complex heterogenous data 
            \item target display spaces
        \end{itemize}
        \pause
        \item structure preserving maps from data to visual
            \begin{itemize}
                \item data and graphics have equivalent continuity
                \item properties are equivariant under monoid actions
            \end{itemize}  
        \pause
        \item composable into complex visualizations
    \end{itemize}
\end{frame}

\begin{frame}[fragile]{How do we make things with this model?}
    \begin{figure}[H]
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/line_1.png}
        \end{subfigure}
    \end{figure}
    \begin{columns}
    \column{0.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Point(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \column{.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Line(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \end{columns}
\end{frame}

\subsection{Visual Transforms}
\begin{frame}[fragile]{How do we implement \vchannel?}
    \begin{minted}{python}
    cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
    transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
                    'y': {'name': 'v2', 'encoder': lambda x: x},
                    'facecolors': {'name':'v3', 'encoder': cmap}, 
                    's':{'name': None , 
                        'encoder': lambda _: itertools.repeat(.02)}}
    \end{minted}
    \begin{itemize}
        \item \mintinline{python}{lambda x: x} is identity \vchannel\
        \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
        \item \mintinline{python}{color.Categorical} is custom \vchannel
    \end{itemize}
\end{frame}

\subsection{Artist}
\begin{frame}[fragile]{Artist} %% rewrite w/ letters in talk
    \begin{minted}{python}
        class ArtistClass(matplotlib.artist.Artist):
            def __init__(self, E, V, *args, **kwargs):
                # set properties that are specific to the artist
                # stash the input E and V
                super().__init__(*args, **kwargs)

            def qhat(self, **args):
                # set the properties of the graphic

            def draw(self, renderer):
                # returns tau, indexed on fiber then key 
                tau = self.E.view(self.axes) 
                # visual channel encoding applied fiberwise 
                visual = {p_i: nu_i(tau_i)
                          for p_i, nu_i, tau)i 
                          in zip(self.V, tau)}
                self.qhat(**visual)
                # pass configurations off to the renderer
                super().draw(renderer)
        \end{minted}
\end{frame}

\begin{frame}[fragile]{Artists: Scatter \& Line}
%%keep the signatures, comments, change names to math
\begin{minted}{python}
    class Point(mcollections.Collection):
        def assemble(self, x, y, s, facecolors='C0' ):
            # construct geometries of the circle glyphs in visual coordinates
            self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
                        for (xi, yi, si) in zip(x, y, s)] 
            # set attributes of glyphs, these are vectorized 
            # circles and facecolors are lists of the same size
            self.set_facecolors(facecolors)

    class Line(mcollections.LineCollection):
        def assemble(self, x, y, color='C0'):
            # assemble line marks as set of segments 
            segments = [np.vstack((vx, vy)).T for vx, vy in zip(x, y)]
            self.set_segments(segments)
            self.set_color(color)
    \end{minted}
\end{frame}

\subsection{Data}
\begin{frame}[fragile]{Continuity}
    \begin{minted}{python}
class PointData: 
    # Fiberbundle is consistent across all sections
    FB = FiberBundle({'tables': ['vertex']},  
        {'v1': float, 'v2': str, 'v3': float})
    def tau(self, k):
        return # tau evaluated at one point k

class LineData: 
    FB = FiberBundle({'tables': ['edge']},
                {'x' : float, 'y':  float, 'color':mtypes.Color()})
    def tau(self, k): 
        return # tau evaluated on interval k
    \end{minted}
\end{frame}
